---
title: "ADA-PARC Category Scorecard"
params:
  category: "WE"   # "CL", "CP", or "WE"
  year: 2024
format:
  html:
    self-contained: true
    toc: false
    page-layout: full
    theme: cosmo
    css: scorecard_v3.css
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(readr)
library(dplyr)
library(ggplot2)
library(gt)
library(maps)
library(scales)
library(htmltools)

# ═══════════════════════════════════════════════════════════════════════════
# CATEGORY CONFIG
# All category-specific values live here. To add a new category, add a
# new entry to this list — the rest of the document is fully generic.
#
# sub_cols_pending: names (from sub_cols) that currently have no data.
#   Set to character(0) when all sub-indices are populated.
# ═══════════════════════════════════════════════════════════════════════════
CATS <- list(

  CL = list(
    name        = "Community Living",
    desc        = "The degree to which people with disabilities (aged 18–64) live in community settings rather than nursing homes, correctional facilities, or other institutions, and whether states invest in community-based housing and support resources.",
    col_overall = "index_rel_community_living",
    col_sub     = c("index_rel_living_arrangements",
                    "index_rel_community_resource"),
    score_col   = "CL Score",
    sub_cols    = c("Living Arrangements", "Community Resource"),
    sub_labels  = c("Living Arrangements",
                    "Community Resource"),
    sub_cols_pending = character(0),
    sub_desc    = c(
      "Composite of four living situation indicators for people with disabilities (aged 18\u201364): share living at home (private residence), in non-institutional group quarters, in nursing homes, and in correctional facilities. Higher scores reflect greater community integration.",
      "Measures state investment in community-based supports: (1) ratio of Medicaid HCBS to institutional spending, (2) HCBS spending per disabled Medicaid enrollee, (3) Housing Choice Voucher households per head-of-household, and (4) Public Housing households per head-of-household."
    ),
    # 4-stop palette: dark red (low/worst) → gold (high/best)
    pal         = c("#7a0901", "#c41306", "#fa7b3c", "#fcde6f"),
    accent      = "#6a040f",
    col_hdr_bg  = "#5a0a0a",
    heading_color = "#92620a",   # section heading color on white bg
    banner_dark = "#92620a",
    banner_light= "#92620a"
  ),

  CP = list(
    name        = "Community Participation",
    desc        = "Access to the tools and systems that enable full participation in community life — including technology, health insurance, educational attainment, transportation, and public safety.",
    col_overall = "index_rel_community_participation",
    col_sub     = c("index_rel_tech",
                    "index_rel_education",
                    "index_rel_insurance",
                    "index_rel_commute",
                    "index_rel_safety"),
    score_col   = "CP Score",
    sub_cols    = c("Technology", "Education", "Insurance",
                    "Transportation", "Safety"),
    sub_labels  = c("Technology Access",
                    "Education Access",
                    "Insurance Access",
                    "Transportation Access",
                    "Public Safety"),
    sub_cols_pending = character(0),
    sub_desc    = c(
      "Computer ownership and broadband internet access among households with disabilities — increasingly essential for employment, healthcare navigation, and civic engagement.",
      "Educational attainment among adults with disabilities, including high school completion and higher education rates — key drivers of economic and civic participation.",
      "Health insurance coverage among people with disabilities aged 19\u201364: private insurance, public insurance (Medicare/Medicaid/CHIP), and uninsured rate.",
      "Commute and transportation accessibility for people with disabilities, reflecting ease of access to employment, services, and community resources.",
      "State-level property and violent crime rates per 100,000 population. Note: FL, PA, and NY have lower agency participation rates."
    ),
    pal         = c("#3b0707", "#780c0c", "#fa7b3c", "#fcd86d"),
    accent      = "#6a040f",
    col_hdr_bg  = "#5a0a0a",
    heading_color = "#92400e",   # section heading color on white bg
    banner_dark = "#c2410c",
    banner_light= "#c2410c"
  ),

  WE = list(
    name        = "Work & Economic",
    desc        = "Economic inclusion and security for people with disabilities — measuring work outcomes (employment, income, poverty, and labor force participation) and housing affordability.",
    col_overall = "index_rel_work_economic",
    col_sub     = c("index_rel_positive_work",
                    "index_rel_negative_work",
                    "index_rel_housing_affordability"),
    score_col   = "WE Score",
    sub_cols    = c("Positive Work", "Negative Work", "Housing Affordability"),
    sub_labels  = c("Positive Work Score",
                    "Negative Work Score",
                    "Housing Affordability"),
    sub_cols_pending = character(0),
    sub_desc    = c(
      "Employment rate and cost-of-living-adjusted income (individual and household) for people with disabilities. Income adjusted using BEA Regional Price Parities (2022).",
      "Negative work indicators for people with disabilities: not-in-labor-force rate, poverty rate (all ages, derived from age-split ACS counts), and unemployment rate. Higher scores indicate lower rates of these negative outcomes.",
      "SSI as a percent of fair market rent for a one-bedroom apartment, plus disability-specific renter and owner cost burden rates. Note: Puerto Rico missing PUMS data."
    ),
    pal         = c("#3b0707", "#780c0c", "#fa7b3c", "#fcd86d"),
    accent      = "#6a040f",
    col_hdr_bg  = "#5a0a0a",
    heading_color = "#6a040f",   # section heading color on white bg
    banner_dark = "#2a0000",
    banner_light= "#3b0707"
  )
)

# ── Select config for this render ──────────────────────────────────────────
cfg <- CATS[[params$category]]

# ═══════════════════════════════════════════════════════════════════════════
# DATA
# ═══════════════════════════════════════════════════════════════════════════
raw     <- read_csv("scorecard_data/final/index_scores_wide.csv",
                    show_col_types = FALSE)
nat_row <- raw |> filter(ABBR == "USA")
states  <- raw |> filter(ABBR != "USA")

# Build scored data frame with generic column names
# Only keep states where the domain-level score is non-NA
scores_df <- states |>
  select(GEOID, NAME, ABBR,
         cat_overall = all_of(cfg$col_overall),
         all_of(cfg$col_sub)) |>
  mutate(cat_overall = as.numeric(cat_overall),
         across(all_of(cfg$col_sub), as.numeric)) |>
  filter(!is.na(cat_overall)) |>
  rename_with(~ paste0("sub_", seq_along(cfg$col_sub)),
              all_of(cfg$col_sub)) |>
  arrange(desc(cat_overall)) |>
  mutate(rank = row_number())

# ── Availability guard ─────────────────────────────────────────────────────
# If ALL domain scores are NA (e.g., WE when sub-index data are pending),
# flag it — remaining chunks will show a pending message and exit early.
data_available <- nrow(scores_df) > 0

# ── Helpers ────────────────────────────────────────────────────────────────
f <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.1f", as.numeric(x)))

# ── Only compute these when data are available ─────────────────────────────
if (data_available) {

  # National values
  nat_overall <- as.numeric(nat_row[[cfg$col_overall]])
  nat_subs    <- sapply(cfg$col_sub, function(x) as.numeric(nat_row[[x]]))

  # Summary stats for banner pills
  top_row <- scores_df |> slice(1)
  bot_row <- scores_df |> slice(n())
  n_above <- sum(scores_df$cat_overall > nat_overall, na.rm = TRUE)

  # ── Quantile breaks: 4 equal-population bins ─────────────────────────────
  q_breaks    <- quantile(scores_df$cat_overall, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  bin_breaks  <- c(0, unname(q_breaks), 100)

  score_pal_4 <- col_bin(
    palette  = cfg$pal,
    bins     = bin_breaks,
    na.color = "#eeeeee"
  )

  # ── Map data ───────────────────────────────────────────────────────────────
  us_map <- map_data("state") |>
    left_join(
      scores_df |> mutate(region = tolower(NAME)),
      by = "region"
    ) |>
    mutate(fill_col = ifelse(is.na(cat_overall), "#d0d5dc",
                             score_pal_4(cat_overall)))

  # ── State label centroids ──────────────────────────────────────────────────
  state_centroids <- us_map |>
    group_by(region) |>
    summarise(long = mean(range(long)), lat = mean(range(lat)), .groups = "drop")

  map_labels <- scores_df |>
    mutate(region = tolower(NAME)) |>
    inner_join(state_centroids, by = "region")

  # ── Census region lookup ───────────────────────────────────────────────────
  census_regions <- c(
    "Connecticut"="Northeast","Maine"="Northeast","Massachusetts"="Northeast",
    "New Hampshire"="Northeast","Rhode Island"="Northeast","Vermont"="Northeast",
    "New Jersey"="Northeast","New York"="Northeast","Pennsylvania"="Northeast",
    "Illinois"="Midwest","Indiana"="Midwest","Iowa"="Midwest",
    "Kansas"="Midwest","Michigan"="Midwest","Minnesota"="Midwest",
    "Missouri"="Midwest","Nebraska"="Midwest","North Dakota"="Midwest",
    "Ohio"="Midwest","South Dakota"="Midwest","Wisconsin"="Midwest",
    "Alabama"="South","Arkansas"="South","Delaware"="South",
    "District of Columbia"="South","Florida"="South","Georgia"="South",
    "Kentucky"="South","Louisiana"="South","Maryland"="South",
    "Mississippi"="South","North Carolina"="South","Oklahoma"="South",
    "South Carolina"="South","Tennessee"="South","Texas"="South",
    "Virginia"="South","West Virginia"="South",
    "Alaska"="West","Arizona"="West","California"="West",
    "Colorado"="West","Hawaii"="West","Idaho"="West",
    "Montana"="West","Nevada"="West","New Mexico"="West",
    "Oregon"="West","Utah"="West","Washington"="West","Wyoming"="West"
  )

  region_df <- scores_df |>
    mutate(region = census_regions[NAME]) |>
    filter(!is.na(region)) |>
    group_by(region) |>
    summarise(avg_score = mean(cat_overall, na.rm = TRUE),
              n_states  = n(), .groups = "drop") |>
    arrange(avg_score) |>
    mutate(region = factor(region, levels = region))

  # ── Top / bottom performers ────────────────────────────────────────────────
  top5 <- scores_df |> slice(1:min(5, n()))
  bot5 <- scores_df |> slice(max(1, n() - 4):n())

}

# ── Inline style strings ───────────────────────────────────────────────────
banner_style  <- paste0(
  "background: ", cfg$banner_light, "; ",
  "color: white; padding: 2.4rem 3rem 2.2rem; ",
  "border-radius: 12px; margin-bottom: 2rem;"
)

heading_style <- paste0(
  "font-family: 'EB Garamond', Garamond, Georgia, serif; ",
  "font-size: 1.6rem; font-weight: 700; color: ", cfg$heading_color, "; ",
  "margin: 0 0 0.3rem; letter-spacing: -0.01em;"
)

heading_style_black <- paste0(
  "font-family: 'EB Garamond', Garamond, Georgia, serif; ",
  "font-size: 1.6rem; font-weight: 700; color: #1c2b3a; ",
  "margin: 0 0 0.3rem; letter-spacing: -0.01em;"
)

# ── Performer card helpers ─────────────────────────────────────────────────
make_perf_row <- function(r, name, score) {
  paste0(
    '<div class="perf-row">',
    '<span class="perf-rank">#', r, '</span>',
    '<span class="perf-name">', name, '</span>',
    '<span class="perf-score">', sprintf("%.1f", score), '</span>',
    '</div>'
  )
}

make_perf_card <- function(label, df, hdr_color) {
  rows <- paste0(mapply(make_perf_row, df$rank, df$NAME, df$cat_overall),
                 collapse = "")
  paste0('<div class="perf-card">',
         '<div class="perf-header" style="background:', hdr_color, ';">', label, '</div>',
         '<div class="perf-body">', rows, '</div>',
         '</div>')
}
```

::: {.page-wrap}

<!-- ═══════════════════════════════════════════════════════════
     BANNER
════════════════════════════════════════════════════════════ -->

<div style="`r banner_style`">

<p class="cat-eyebrow">ADA-PARC Scorecard &nbsp;·&nbsp; `r params$year` Data</p>

<h1 style="font-family:'EB Garamond',Garamond,Georgia,serif; font-size:3rem; font-weight:700; color:white; margin:0 0 0.3rem; line-height:1.05;">`r cfg$name`</h1>

<p class="cat-desc">`r cfg$desc`</p>

```{r banner-pills, results='asis'}
if (!data_available) {
  cat(paste0(
    '<div style="background:rgba(255,255,255,0.15); border-radius:10px;',
    ' padding:1.2rem 1.8rem; margin-top:1rem; color:white;">',
    '<strong>Domain scores are not available for this category.</strong><br>',
    'All sub-index scores are NA for the selected year.',
    ' Re-run the pipeline to regenerate scores, or check that source data',
    ' files are present in the expected locations.',
    '</div>'
  ))
} else {
  cat(paste0(
    '::: {.stat-pills}\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', f(nat_overall), '</div>\n',
    '<div class="pill-lbl">US Average</div>\n',
    ':::\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', f(top_row$cat_overall), '</div>\n',
    '<div class="pill-lbl">Highest Score</div>\n',
    '<div class="pill-sub">', top_row$NAME, '</div>\n',
    ':::\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', f(bot_row$cat_overall), '</div>\n',
    '<div class="pill-lbl">Lowest Score</div>\n',
    '<div class="pill-sub">', bot_row$NAME, '</div>\n',
    ':::\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', n_above, '</div>\n',
    '<div class="pill-lbl">States Above US Avg</div>\n',
    '<div class="pill-sub">of ', nrow(scores_df), ' states &amp; territories</div>\n',
    ':::\n\n',
    ':::\n'
  ))
}
```

</div>

```{r early-exit, results='asis'}
# If no data, show an explanatory pending section and stop rendering
if (!data_available) {
  cat(paste0(
    '<div class="section-card" style="margin-top:1.5rem;">',
    '<p style="', heading_style, '">About This Category</p>',
    '<p class="section-sub">The sub-indices that comprise the ',
    cfg$name, ' score and the measures they capture.</p>',
    '<ul class="about-list">',
    paste0(
      "<li><strong>", cfg$sub_labels, "</strong> &mdash; ", cfg$sub_desc, "</li>",
      collapse = "\n"
    ),
    '</ul>',
    '</div>'
  ))
  knitr::knit_exit()
}
```

<!-- ═══════════════════════════════════════════════════════════
     MAP
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<p style="`r heading_style`">State Scores — `r cfg$name`</p>
<p class="section-sub">States divided into four equal-population quartile groups. Alaska, Hawaii, D.C., and territories appear in the rankings table below.</p>

```{r choropleth, fig.width=10, fig.height=6.2, fig.align='center'}
# Build legend labels from quartile breaks
legend_labels <- c(
  paste0("Q1: 0\u2013",    sprintf("%.1f", q_breaks[1])),
  paste0("Q2: ", sprintf("%.1f", q_breaks[1]), "\u2013", sprintf("%.1f", q_breaks[2])),
  paste0("Q3: ", sprintf("%.1f", q_breaks[2]), "\u2013", sprintf("%.1f", q_breaks[3])),
  paste0("Q4: ", sprintf("%.1f", q_breaks[3]), "\u2013100")
)

ggplot(us_map, aes(x = long, y = lat, group = group, fill = fill_col)) +
  geom_polygon(color = "white", linewidth = 0.35) +
  scale_fill_identity(
    name   = "Score (quartile)",
    guide  = guide_legend(
      title.position = "top",
      title.hjust    = 0.5,
      keywidth       = unit(1.8, "cm"),
      keyheight      = unit(0.45, "cm"),
      label.theme    = element_text(size = 9, color = "#1c2b3a", family = "serif")
    ),
    breaks = cfg$pal,
    labels = legend_labels
  ) +
  geom_text(
    data = map_labels,
    aes(x = long, y = lat, label = ABBR),
    inherit.aes = FALSE,
    size = 2.5, color = "#1c2b3a", fontface = "bold", alpha = 0.75
  ) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  labs(x = NULL, y = NULL) +
  theme_void(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.title    = element_text(size = 10, face = "bold",
                                   color = "#1c2b3a", family = "serif"),
    legend.text     = element_text(size = 9, color = "#1c2b3a",
                                   family = "serif"),
    plot.background  = element_blank(),
    panel.background = element_blank()
  )
```

<p class="map-caption">Scores range 0–100; higher is better. Grey indicates states with no data in this render.</p>

```{r strip-chart, fig.width=10, fig.height=2.6, fig.align='center'}
strip_df <- scores_df |>
  select(NAME, ABBR, cat_overall) |>
  mutate(fill_col = score_pal_4(cat_overall))

ggplot(strip_df, aes(x = cat_overall, y = 0.5)) +
  # Quartile region shading
  annotate("rect", xmin = 0,              xmax = q_breaks[[1]],
           ymin = 0, ymax = 1, fill = cfg$pal[1], alpha = 0.13) +
  annotate("rect", xmin = q_breaks[[1]],  xmax = q_breaks[[2]],
           ymin = 0, ymax = 1, fill = cfg$pal[2], alpha = 0.13) +
  annotate("rect", xmin = q_breaks[[2]],  xmax = q_breaks[[3]],
           ymin = 0, ymax = 1, fill = cfg$pal[3], alpha = 0.13) +
  annotate("rect", xmin = q_breaks[[3]],  xmax = 100,
           ymin = 0, ymax = 1, fill = cfg$pal[4], alpha = 0.13) +
  # Quartile break lines
  geom_vline(xintercept = unname(q_breaks),
             color = "#cccccc", linewidth = 0.5) +
  # National average dashed line
  geom_vline(xintercept = nat_overall,
             color = "#1c2b3a", linewidth = 0.85, linetype = "dashed") +
  # State dots, colored by palette bin
  geom_jitter(aes(color = fill_col),
              position = position_jitter(height = 0.26, width = 0, seed = 42),
              size = 3.4, alpha = 0.88) +
  scale_color_identity() +
  # US average label
  annotate("text", x = nat_overall + 1.5, y = 0.87,
           label = paste0("US avg  ", f(nat_overall)),
           hjust = 0, size = 3, color = "#1c2b3a",
           fontface = "bold", family = "serif") +
  # Quartile labels
  annotate("text", x = q_breaks[[1]] / 2,                    y = 0.1,
           label = "Q1", size = 2.4, color = "#aaaaaa", family = "serif") +
  annotate("text", x = (q_breaks[[1]] + q_breaks[[2]]) / 2,  y = 0.1,
           label = "Q2", size = 2.4, color = "#aaaaaa", family = "serif") +
  annotate("text", x = (q_breaks[[2]] + q_breaks[[3]]) / 2,  y = 0.1,
           label = "Q3", size = 2.4, color = "#aaaaaa", family = "serif") +
  annotate("text", x = (q_breaks[[3]] + 100) / 2,            y = 0.1,
           label = "Q4", size = 2.4, color = "#aaaaaa", family = "serif") +
  scale_x_continuous(limits = c(0, 100),
                     breaks = c(0, 25, 50, 75, 100),
                     expand = c(0.01, 0)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  labs(x = "Score (0\u2013100)", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(color = "#eaecf0", linewidth = 0.3),
    axis.text.x        = element_text(size = 9, color = "#555555", family = "serif"),
    axis.title.x       = element_text(size = 9, color = "#555555", family = "serif",
                                      margin = margin(t = 5)),
    plot.background    = element_blank(),
    panel.background   = element_blank(),
    plot.margin        = margin(10, 20, 2, 20)
  )
```

:::

<!-- ═══════════════════════════════════════════════════════════
     RANKINGS TABLE
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<p style="`r heading_style_black`">State Rankings</p>
<p class="section-sub">All states, D.C., and Puerto Rico ranked by `r cfg$name` score. Sub-index scores are shown alongside the overall score.</p>

```{r gt-table}
# ── Build display table with pretty column names ───────────────────────────
sub_rename <- setNames(paste0("sub_", seq_along(cfg$sub_cols)), cfg$sub_cols)

tbl_data <- scores_df |>
  mutate(Tier = dplyr::case_when(
    rank <= 12 ~ "Excellent",
    rank <= 25 ~ "Above Avg",
    rank <= 39 ~ "Below Avg",
    TRUE       ~ "Poor"
  )) |>
  rename(!!cfg$score_col := cat_overall) |>
  rename(!!!sub_rename) |>
  select(rank, Tier, State = NAME, Abbr = ABBR,
         all_of(cfg$score_col),
         all_of(cfg$sub_cols))

# ── All score columns for fmt_number ──────────────────────────────────────
score_cols <- c(cfg$score_col, cfg$sub_cols)

# ── Build source note with national averages (NA-safe) ────────────────────
nat_note_parts <- paste0(
  f(nat_subs), " ", cfg$sub_cols
)
nat_note <- paste0(
  "**US National Average:** ",
  f(nat_overall), " overall",
  paste0(" &nbsp;\u00b7&nbsp; ", nat_note_parts, collapse = "")
)

# ── Render gt table ────────────────────────────────────────────────────────
tbl_data |>
  gt() |>

  fmt_number(columns = all_of(score_cols), decimals = 1) |>

  # Show — for NA values in sub-index columns
  fmt_missing(columns = all_of(score_cols), missing_text = "\u2014") |>

  cols_label(rank = "Rank", Tier = "Tier", Abbr = "",
             State = "State / Territory") |>

  cols_width(
    rank  ~ px(55),
    Tier  ~ px(95),
    Abbr  ~ px(48),
    State ~ px(190),
    everything() ~ px(105)
  ) |>

  cols_align(align = "center",
             columns = c(rank, Tier, Abbr, all_of(cfg$score_col),
                         all_of(cfg$sub_cols))) |>
  cols_align(align = "left", columns = State) |>

  # Colour overall score column only
  data_color(
    columns        = all_of(cfg$score_col),
    fn             = score_pal_4,
    autocolor_text = TRUE
  ) |>

  # Bold rank
  tab_style(
    style     = cell_text(weight = "bold", size = px(15)),
    locations = cells_body(columns = rank)
  ) |>

  # Accent border on top 3 rows
  tab_style(
    style     = cell_borders(sides = "left", color = cfg$accent, weight = px(4)),
    locations = cells_body(rows = rank <= 3)
  ) |>

  tab_source_note(source_note = md(nat_note)) |>

  tab_header(
    title    = md(paste0("**", cfg$name, "** — All States Ranked")),
    subtitle = md(paste0(
      "Ranked 1–", nrow(tbl_data),
      " by overall ", cfg$name, " score · ",
      params$year, " ACS Data"
    ))
  ) |>

  tab_options(
    table.width                     = pct(100),
    table.font.names                = c("EB Garamond", "Georgia", "serif"),
    table.font.size                 = px(16),
    data_row.padding                = px(5),
    heading.background.color        = cfg$accent,
    heading.title.font.size         = px(18),
    heading.subtitle.font.size      = px(13),
    column_labels.background.color  = cfg$col_hdr_bg,
    column_labels.font.size         = px(13),
    row.striping.include_table_body = TRUE,
    row.striping.background_color   = "#f8f8fc",
    source_notes.font.size          = px(13),
    source_notes.background.color   = "#f8f8fc",
    table.border.top.color          = cfg$accent,
    table.border.top.width          = px(3)
  ) |>

  tab_style(
    style     = cell_text(color = "white", weight = "bold"),
    locations = cells_column_labels(everything())
  ) |>
  tab_style(
    style     = cell_text(color = "white"),
    locations = cells_title(groups = c("title", "subtitle"))
  )
```

:::

<!-- ═══════════════════════════════════════════════════════════
     TOP / BOTTOM PERFORMERS
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<p style="`r heading_style`">Top &amp; Bottom Performers</p>
<p class="section-sub">The five highest- and lowest-scoring states for `r cfg$name`.</p>

```{r performers}
HTML(paste0(
  '<div class="top-bottom-container">',
  make_perf_card("Top 5 States",    top5, "#92620a"),
  make_perf_card("Bottom 5 States", bot5, "#7b1717"),
  '</div>'
))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     REGIONAL BREAKDOWN
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<p style="`r heading_style`">Average Score by Census Region</p>
<p class="section-sub">State scores averaged within each U.S. Census region. Territories and Puerto Rico excluded.</p>

```{r regional-chart, fig.width=8, fig.height=2.4, fig.align='center'}
ggplot(region_df, aes(x = avg_score,
                      y = region)) +
  geom_col(fill = cfg$pal[3], alpha = 0.82, width = 0.48) +
  geom_vline(xintercept = nat_overall,
             linetype = "dashed", color = "#1c2b3a", linewidth = 0.75) +
  geom_text(aes(label = sprintf("%.1f", avg_score)),
            hjust = -0.25, size = 3.8, color = "#1c2b3a",
            family = "serif", fontface = "bold") +
  annotate("text", x = nat_overall + 0.8, y = 0.55,
           label = paste0("US avg  ", f(nat_overall)),
           hjust = 0, size = 3, color = "#1c2b3a",
           fontface = "bold", family = "serif") +
  scale_x_continuous(limits = c(0, 100),
                     breaks = c(0, 25, 50, 75, 100),
                     expand = c(0.01, 0)) +
  labs(x = "Average Score (0\u2013100)", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y        = element_text(size = 12, family = "serif",
                                      color = "#1c2b3a"),
    axis.text.x        = element_text(size = 9, color = "#555555"),
    axis.title.x       = element_text(size = 9, color = "#555555",
                                      margin = margin(t = 5)),
    panel.grid.major.x = element_line(color = "#eaecf0", linewidth = 0.4),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.background    = element_blank(),
    panel.background   = element_blank(),
    plot.margin        = margin(4, 20, 4, 4)
  )
```

:::

<!-- ═══════════════════════════════════════════════════════════
     ABOUT THIS CATEGORY
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<p style="`r heading_style`">About This Category</p>
<p class="section-sub">The sub-indices that comprise the `r cfg$name` score and the measures they capture.</p>

```{r about-measures, results='asis'}
items <- paste0(
  "<li><strong>", cfg$sub_labels, "</strong> &mdash; ", cfg$sub_desc, "</li>",
  collapse = "\n"
)
cat(paste0('<ul class="about-list">', items, '</ul>'))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     FOOTER
════════════════════════════════════════════════════════════ -->
<p class="footer-note">
  <strong>Data source:</strong> ADA-PARC Scorecard &nbsp;·&nbsp;
  `r params$year` American Community Survey (ACS) &nbsp;·&nbsp;
  Scores range 0–100; higher is better. &nbsp;·&nbsp;
  Sub-indices comprising the <strong>`r cfg$name`</strong> score:
  `r paste(cfg$sub_labels, collapse = "; ")`.
  Alaska, Hawaii, D.C., and Puerto Rico are excluded from the map but included in the rankings table.
</p>

:::
